---
title: "World Happiness"
subtitle: "2022"
author: "Ben Mitchell, Eddie Jiang, Callum Knox, Sarah Brown-Scott"
institute: "University of Edinburgh"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r load-packages, include = FALSE}
# Add any additional packages you need to this chunk
library(tidyverse)
library(palmerpenguins)
library(knitr)
library(xaringanthemer)
library(tidymodels)
library(broom)
```

```{r setup, include=FALSE}
# For better figure resolution
knitr::opts_chunk$set(fig.retina = 3, dpi = 300, fig.width = 6, fig.asp = 0.618, out.width = "80%")
```

```{r include=FALSE}
style_xaringan(
  title_slide_background_image = "images_for_project/yellow_backround.jpg"
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
style_mono_dark(base_color = "ivory")
```

```{r load-data, include=FALSE}
# Load your data here
happiness <- read_csv("presentation_files/2015.csv")
countries <- read_csv("presentation_files/countries of the world.csv")
```

```{r include=FALSE}
countries$Country[countries$Country == "Cote d\'Ivoire"] = "Ivory Coast"
countries$Country[countries$Country == "Bosnia & Herzegovina"] = "Bosnia and Herzegovina"
countries$Country[countries$Country == "Central African Rep."] = "Central African Republic"
countries$Country[countries$Country == "Congo, Dem. Rep."] = "Congo (Kinshasa)"
countries$Country[countries$Country == "Congo, Repub. of the"] = "Congo (Brazzaville)"
# error in kososvo
# and montenegro cont
```

```{r include=FALSE}
happiness_countries <- left_join(happiness, countries, by = "Country")
```

```{r include=FALSE}
colnames(happiness_countries)[4] <- "happiness_score"
colnames(happiness_countries)[6] <- "economy"
colnames(happiness_countries)[7] <- "family"
colnames(happiness_countries)[8] <- "health"
colnames(happiness_countries)[10] <- "trust"
colnames(happiness_countries)[9] <- "freedom"
colnames(happiness_countries)[20] <- "gdp"
colnames(happiness_countries)[19] <- "infant_mortality"
colnames(happiness_countries)[22] <- "phones"
```

```{r include=FALSE, eval = FALSE}
#"we wanted to see if a countries gdp actually affected how happy people were with the economy, as the happiness score for economy is based on perception."

#model predicting economy score from GDP
set.seed(1234)

economy_boot <- bootstraps(happiness_countries, times = 5000)

economy_models <- economy_boot %>%
  mutate(
    model = map(splits, ~lm(gdp ~ economy, data = .)),
    coef_info = map(model, tidy)
  )

economy_coef <- economy_models %>%
  unnest(coef_info)

economy_confidence <- int_pctl(economy_models, coef_info)

#we are 95% confident that for each additional  0.1 of a point of economy score, we would expect gdp to increase by $1734.00 to $2390.46, on average.
```

```{r include=FALSE, eval = FALSE}
#"we wanted to see if a countries infant mortality actually affected how happy people were with the healthcare, as the happiness score for health is based on perception."

#model predicting health score from infant mortality
set.seed(1234)

health_boot <- bootstraps(happiness_countries, times = 5000)

health_models <- health_boot %>%
  mutate(
    model = map(splits, ~lm(infant_mortality ~ health, data = .)),
    coef_info = map(model, tidy)
  )

haelth_coef <- health_models %>%
  unnest(coef_info)

health_confidence <- int_pctl(health_models, coef_info)

#we are 95% confident that for each additional  0.1 of a point of health score, we would expect infant mortality to decrease by 9637.09 to 14180.66, on average.
```

```{r include=TRUE, eval = TRUE}
happiness_countries %>%
    drop_na(infant_mortality) %>%
  ggplot(data = happiness_countries,
         mapping = aes(x = health, y = infant_mortality,
                       na.rm = TRUE))+
  geom_jitter() +
  geom_smooth(method = "lm", se = F) +
  labs(title = "Infant Mortality against Health",
       x = "Health Score",
       y = "Infant Mortality") +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        plot.background = element_rect(fill = "aquamarine4"),
        axis.line = element_line(size = rel(3), 
                                 arrow = arrow(), 
                                 colour = "black"),
        panel.background = element_rect(fill = "gray90", 
                                        colour = "aquamarine4"),
        axis.title.y = element_text(size = rel(1.5), 
                                    angle = (90),
                                    color = "white"),
        axis.title.x = element_text(size = rel(1.5),
                                    color = "white"),
        panel.grid.major = element_line(colour = "aquamarine4"),
        panel.grid = element_line(colour = "aquamarine4"),
        plot.title = element_text(size = rel(1.2),
                                  color = "white"),
        legend.text = element_text(size = rel(1.5),
                                   color = "black"),
        legend.title = element_text(size = rel(1.5),
                                   color = "black"),
        legend.background = element_rect(fill = "gray90"),
        legend.key = element_rect(fill = "gray90"),
        axis.text = element_text(color = "white")
        )
```

```{r include=FALSE, eval = FALSE}
#"we wanted to see if a countries phones per 1000 people actually affected how happy people were."

#model predicting happiness from phones per 1000
set.seed(7399)

phones_boot <- bootstraps(happiness_countries, times = 5000)

phones_models <- phones_boot %>%
  mutate(
    model = map(splits, ~lm(happiness_score ~ phones, data = .)),
    coef_info = map(model, tidy)
  )

phones_coef <- phones_models %>%
  unnest(coef_info)

phones_confidence <- int_pctl(phones_models, coef_info)

#we are 95% confident that for each additional  1 phones per 1000 people, we would expect the happiness score to increase by  0.90 to 2.56, on average.
```


```{r include=FALSE, eval = FALSE}
phones_confidence_mean <- phones_confidence %>%
  filter(term != "(Intercept)")

lower <- mean(phones_confidence_0$.lower)
upper <- mean(phones_confidence_0$.upper)
```

```{r include=TRUE, eval = TRUE}
happiness_countries %>%
  drop_na(phones) %>%
  ggplot(data = happiness_countries,
         mapping = aes(x = phones, y = happiness_score,
                       na.rm = TRUE))+
  geom_jitter() +
  expand_limits(y = c(0, 10)) +
  labs(title = "Happiness Score against Phones per 1000 People",
       x = "Phones Per 1000 People",
       y = "Happiness Score") +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        plot.background = element_rect(fill = "aquamarine4"),
        axis.line = element_line(size = rel(3), 
                                 arrow = arrow(), 
                                 colour = "black"),
        panel.background = element_rect(fill = "gray90", 
                                        colour = "aquamarine4"),
        axis.title.y = element_text(size = rel(1.5), 
                                    angle = (90),
                                    color = "white"),
        axis.title.x = element_text(size = rel(1.5),
                                    color = "white"),
        panel.grid.major = element_line(colour = "aquamarine4"),
        panel.grid = element_line(colour = "aquamarine4"),
        plot.title = element_text(size = rel(1.2),
                                  color = "white"),
        legend.text = element_text(size = rel(1.5),
                                   color = "black"),
        legend.title = element_text(size = rel(1.5),
                                   color = "black"),
        legend.background = element_rect(fill = "gray90"),
        legend.key = element_rect(fill = "gray90"),
        axis.text = element_text(color = "white")
        )
```

class: center, middle

## How  Do a Range of Factors Affect Worldwide Happiness???


---

class: center, middle

# Dataset Information

-To try and answer this question we picked 2 data sets.

-They were "World happiness Report" from 2015, and "Countries of The World.

-World Happiness Report contains data from a survey about overall happiness in 158 countries

```{r echo=FALSE}
glimpse(happiness)
```
- Variables economy, family, health, freedom, trust, and generosity show how happy people are with these things.

---

class: inverse, center, middle

# Reasons Behind Question Choice

---
class: center, middle

- We selected this question as mental health is a prominent issue in society and analysis of the data in this way could give some indicators of ways to improve the overall happiness of a country.


- The variables we think will best suit our question and hypotheses are, government trust, how the country is doing in terms of GDP and economy, the countries citizen freedom rating and the projected life expectancy of its citizens.
 And how these factors can vary independently and jointly with the countries happiness rating. 
If factors are proportional or inverse with each other. 


---

class: inverse, middle, center

# Plots

```{r echo=FALSE}
happiness %>%
ggplot(data = happiness_countries, 
       mapping = aes(x = family, y = happiness_score))+
  geom_jitter() +
  geom_smooth(method = "lm", se = F) +
  labs(title = "Happiness Score with Family",
       x = "Family",
       y = "Happiness Score") + 
  theme(plot.background = element_rect(fill = "aquamarine4"),
        axis.line = element_line(size = rel(3), 
                                 arrow = arrow(), 
                                 colour = "black"),
        panel.background = element_rect(fill = "gray90", 
                                        colour = "aquamarine4"),
        axis.title.y = element_text(size = rel(1.5), 
                                    angle = (90),
                                    color = "white"),
        axis.title.x = element_text(size = rel(1.5),
                                    color = "white"),
        panel.grid.major = element_line(colour = "aquamarine4"),
        panel.grid = element_line(colour = "aquamarine4"),
        plot.title = element_text(size = rel(1.2),
                                  color = "white"),
        legend.text = element_text(size = rel(1.5),
                                   color = "black"),
        legend.title = element_text(size = rel(1.5),
                                   color = "black"),
        legend.background = element_rect(fill = "gray90"),
        legend.key = element_rect(fill = "gray90"),
        axis.text = element_text(color = "white")
        )
```


---

class: inverse, middle, center

# Plots

```{r echo=FALSE}
happiness %>%
ggplot(data = happiness_countries, 
       mapping = aes(x = freedom, y = happiness_score))+
  geom_jitter()+
  geom_smooth(method = "lm", se = F) +
  labs(title = "Happiness Score with Freedom",
       x = "Freedom",
       y = "Happiness Score") + 
  theme(plot.background = element_rect(fill = "aquamarine4"),
        axis.line = element_line(size = rel(3), 
                                 arrow = arrow(), 
                                 colour = "black"),
        panel.background = element_rect(fill = "gray90", 
                                        colour = "aquamarine4"),
        axis.title.y = element_text(size = rel(1.5), 
                                    angle = (90),
                                    color = "white"),
        axis.title.x = element_text(size = rel(1.5),
                                    color = "white"),
        panel.grid.major = element_line(colour = "aquamarine4"),
        panel.grid = element_line(colour = "aquamarine4"),
        plot.title = element_text(size = rel(1.2),
                                  color = "white"),
        legend.text = element_text(size = rel(1.5),
                                   color = "black"),
        legend.title = element_text(size = rel(1.5),
                                   color = "black"),
        legend.background = element_rect(fill = "gray90"),
        legend.key = element_rect(fill = "gray90"),
        axis.text = element_text(color = "white")
        )

```


---
class: inverse, middle, center

# Plots

```{r echo=FALSE}
happiness %>%
ggplot(data = happiness_countries, 
       mapping = aes(x = health, y = happiness_score))+
  geom_jitter()+
  geom_smooth(method = "lm", se = F) +
  labs(title = "Happiness Score with Health",
       x = "Health",
       y = "Happiness Score") + 
  theme(plot.background = element_rect(fill = "aquamarine4"),
        axis.line = element_line(size = rel(3), 
                                 arrow = arrow(), 
                                 colour = "black"),
        panel.background = element_rect(fill = "gray90", 
                                        colour = "aquamarine4"),
        axis.title.y = element_text(size = rel(1.5), 
                                    angle = (90),
                                    color = "white"),
        axis.title.x = element_text(size = rel(1.5),
                                    color = "white"),
        panel.grid.major = element_line(colour = "aquamarine4"),
        panel.grid = element_line(colour = "aquamarine4"),
        plot.title = element_text(size = rel(1.2),
                                  color = "white"),
        legend.text = element_text(size = rel(1.5),
                                   color = "black"),
        legend.title = element_text(size = rel(1.5),
                                   color = "black"),
        legend.background = element_rect(fill = "gray90"),
        legend.key = element_rect(fill = "gray90"),
        axis.text = element_text(color = "white")
        )

```


---
class: inverse, middle, center

# Plots

```{r echo=FALSE}
happiness %>%
ggplot(data = happiness_countries, 
       mapping = aes(x = trust, y = happiness_score))+
  geom_jitter()+
  geom_smooth(method = "lm", se = F) +
  labs(title = "Happiness Score with Government Trust",
       x = "Trust",
       y = "Happiness Score") + 
  theme(plot.background = element_rect(fill = "aquamarine4"),
        axis.line = element_line(size = rel(3), 
                                 arrow = arrow(), 
                                 colour = "black"),
        panel.background = element_rect(fill = "gray90", 
                                        colour = "aquamarine4"),
        axis.title.y = element_text(size = rel(1.5), 
                                    angle = (90),
                                    color = "white"),
        axis.title.x = element_text(size = rel(1.5),
                                    color = "white"),
        panel.grid.major = element_line(colour = "aquamarine4"),
        panel.grid = element_line(colour = "aquamarine4"),
        plot.title = element_text(size = rel(1.2),
                                  color = "white"),
        legend.text = element_text(size = rel(1.5),
                                   color = "black"),
        legend.title = element_text(size = rel(1.5),
                                   color = "black"),
        legend.background = element_rect(fill = "gray90"),
        legend.key = element_rect(fill = "gray90"),
        axis.text = element_text(color = "white")
        )

```


---
class: inverse, middle, center

# Plots

```{r echo=FALSE}
happiness %>%
ggplot(data = happiness_countries, 
       mapping = aes(x = economy, y = happiness_score))+
  geom_jitter()+
  geom_smooth(method = "lm", se = F) +
  labs(title = "Happiness Score with Economy", 
       x = "Economy",
       y = "Happiness Score") + 
  theme(plot.background = element_rect(fill = "aquamarine4"),
        axis.line = element_line(size = rel(3), 
                                 arrow = arrow(), 
                                 colour = "black"),
        panel.background = element_rect(fill = "gray90", 
                                        colour = "aquamarine4"),
        axis.title.y = element_text(size = rel(1.5), 
                                    angle = (90),
                                    color = "white"),
        axis.title.x = element_text(size = rel(1.5),
                                    color = "white"),
        panel.grid.major = element_line(colour = "aquamarine4"),
        panel.grid = element_line(colour = "aquamarine4"),
        plot.title = element_text(size = rel(1.2),
                                  color = "white"),
        legend.text = element_text(size = rel(1.5),
                                   color = "black"),
        legend.title = element_text(size = rel(1.5),
                                   color = "black"),
        legend.background = element_rect(fill = "gray90"),
        legend.key = element_rect(fill = "gray90"),
        axis.text = element_text(color = "white")
        )

```

---
class: inverse, middle, center

# Plots

```{r echo=FALSE}
happiness %>%
  ggplot(data = happiness_countries, 
         mapping = aes(x = economy, y = gdp))+
  geom_jitter()+
  geom_smooth(method = "loess", se = F) +
  labs(title = "Economy Score against GDP",
       x = "Economy Score",
       y = "GDP") + 
  theme(plot.background = element_rect(fill = "aquamarine4"),
        axis.line = element_line(size = rel(3), 
                                 arrow = arrow(), 
                                 colour = "black"),
        panel.background = element_rect(fill = "gray90", 
                                        colour = "aquamarine4"),
        axis.title.y = element_text(size = rel(1.5), 
                                    angle = (90),
                                    color = "white"),
        axis.title.x = element_text(size = rel(1.5),
                                    color = "white"),
        panel.grid.major = element_line(colour = "aquamarine4"),
        panel.grid = element_line(colour = "aquamarine4"),
        plot.title = element_text(size = rel(1.2),
                                  color = "white"),
        legend.text = element_text(size = rel(1.5),
                                   color = "black"),
        legend.title = element_text(size = rel(1.5),
                                   color = "black"),
        legend.background = element_rect(fill = "gray90"),
        legend.key = element_rect(fill = "gray90"),
        axis.text = element_text(color = "white")
        )

```
---
---
class: inverse, middle, center

# Regions Plot

```{r echo=FALSE}
happiness_countries %>%
  ggplot(data = happiness_countries, 
         mapping = aes(x = happiness_score, y = Region.x))+
  geom_boxplot()+
  labs(title = "Happiness Score shown by Region",
       subtitle = "Regions are split into sub regions",
       x = "Happiness Score",
       y = "Region") + 
  theme(plot.background = element_rect(fill = "aquamarine4"),
        axis.line = element_line(size = rel(3), 
                                 arrow = arrow(), 
                                 colour = "black"),
        panel.background = element_rect(fill = "gray90", 
                                        colour = "aquamarine4"),
        axis.title.y = element_text(size = rel(1.5), 
                                    angle = (90),
                                    color = "white"),
        axis.title.x = element_text(size = rel(1.5),
                                    color = "white"),
        panel.grid.major = element_line(colour = "aquamarine4"),
        panel.grid = element_line(colour = "aquamarine4"),
        plot.title = element_text(size = rel(1.2),
                                  color = "white"),
        plot.subtitle = element_text(size = rel(1),
                                  color = "white"),
        legend.text = element_text(size = rel(1.5),
                                   color = "black"),
        legend.title = element_text(size = rel(1.5),
                                   color = "black"),
        legend.background = element_rect(fill = "gray90"),
        legend.key = element_rect(fill = "gray90"),
        axis.text = element_text(color = "white")
        )
```

---
